Election Preferences in 2022 based on Gender and Education

### 2024-04-19 paper/paper.qmd
more ediiting to the qmd document and editted some scripts as well

The data used in this paper is contained in the Cooperative Election Study 2022.@dataset This study examines American's views on representatives, electoral experiences and elections. This data is available for free at the Harvard University Dataverse.

Surverys were used to collect data from a sample large enough that it may be assumed is representative of the entire American population. Importantly, this covered a wide range of constituencies and the sample was large enough such that the data for each constituency is enough to assume that it is representative of its whole.@dataset There were 60 research teams, and each purchased 1000 surverys to hand out to their respective constituencies. Sample matching was used to pick the 1000 people in which the surverys were to be administered. This is a process which is not truly random but has the benefit of being practical. In a truly randomly selected group, the contact information for some may not be readily available and therefore makes the data collectin process much more difficult. Sample matching involves selecting a sample at random, and anyone who does not fit the criteria to proceed(in this case having your contact information available) is then replaced by a person in the pool who is believed to have similar characteristics.

@@

Define $y_i$ as the political preference of the individual and is 1 if the individual prefers Biden and 0 if the person prefers trump. Then $\mbox{gender}_i$ is the individual's gender and $\mbox{education}_i$ is the individual's education

@@

## Female Prefereces

We observe that females are 0.8 times as likely to vote for Trump as they are for Biden. This follows a trend that has been observed since the beginning of the 21st century, women prefer the democratic party.@article1 This may be due to the history of the democratic party and its advocation for women-rights. It appears that the democratic party is the one that serves women best in terms of their economic interest. In fact the democratic party has paid a lot more attention to issues that directly pertain to women such as abortion and family leave than the republicans and there is evidence to suggest that this is a key part of its strategy. @article2

## Highly Educated People

One of the most apparent outcomes of this model would be the likelihood of highly educated people to prefer voting for Biden over Trump. One cause of this can be the fact that college graduates are far more likely to self-identify as liberal.@article3 It could also pertain to the way in which most colleges are structured. It is possible that the values taught there are aligned with those of the democrats and hence shape the thinking of these highly educated people in a way that increases their likelihood of voting for a specific candidate.

The proportion of graduates that make up the democratic voting pool has greatly increased over the last 20 years.@article3

## Political Strategy
The most interesting part is how these observations apply to the political strategy of the two parties. Firstly, it was argued that the majority of females voting for the democrats were enough to sway the elections and therefore the main focus of the two parties, however this was disproven in 1980 and 1984 as the republican party won without fully closing the gender gap. They appealed to broad groups of women through issues that directly affected them and were not specific to women such as national security.@article2

Additionally, with the increased amount of liberal, educated individuals arguing for various policies that inconvenience some working class individuals, it creates an opportunity for the republican party to increase its share of this demographic in an attempt to balance the disproportionality. An exaple of this may be the advocation for an environmentally friendly policy that negatively impacts the working class people in this field. (@article3)

## Weaknesses and next steps

A major weakness of this analysis is that it assumes that Biden and Trump are choices made because of preferences. However, as explained in the model section, these are the only rational choices as any other choice seems to be irrelevant. Therefore someone may vote for a candidate not because they resonate with their political ideals, but rather because they are the 'lesser of two evils.'

One way in which this can be analyzed is by considering someones thoughts on different administrations throughout the years or using a benchmark to determine how much they genuinely believe in the political ideals of a certain party.

>>>>

The data used in this paper was downloaded directly fromt the Riot Games API, the company that owns league of legends. This is available to everyone for free(although an account must be created.) When a game of league of legends is played online, a large amount of data is recorded and stored in their database. These can then be accessed using their API, which allows this data to be queried. The data used in this paper comes from games played by each of the players on the pro teams, and were within the last 2 years. This allows it to be more indicative of their current performance, rather than their performance from 2016. It is taken from their solo queue games, which is essentially the regular games that anyone can play online by themselves and not the professional games in which all of the team's players team up and play against another full team. (As the name suggests, solo queue means you are queueing for a game alone.)

@@

Define $y_i$ as the outcome of the game and is 1 if the player wins the game and 0 if the person loses. Then $\mbox{CS}_i$ is the player's creep score at 10 minutes and $\mbox{Level}_i$ is the player's highest level lead against their lane opponent.

@@

##LEAGUE OF LEGENDS 
###PLAYER ROLES

As mentioned previously, League of Legends has three lanes: top lane, mid lane, and bot lane. Usually, your tank goes in the top lane, which is the player that doesn’t deal as much damage as the rest but takes a long time to kill. The mid lane has a mage who deals a different type of damage than the rest of their team and requires a different type of resistance to be nullified by the enemy team. The bot lane consists of the Carry and the Support. The carry is somewhat of a glass cannon, as they are very easy to kill but usually deal insane amounts of damage if left unchecked. The carry also usually follows a progression of being weak in the early stages of the game and strong in the late stages. This is why it is usually optimal to give the carry another player to support them, the support. The support's job is simple: babysit the carry and play for the team. The support champions are usually equipped with abilities that fulfill one of two roles: engage supports or enchanter supports. Enchanter supports buff the allied carry with a variety of shields, heals, and attack speed boosts. Engage supports have abilities that can lock down the enemy team and position them in such a way that it is easy for the carry to deal damage. The fifth player on the team is always the Jungler. They roam between lanes, and if any of the other four players on the enemy team are not careful, the Jungler can surprise them with a fight in which the advantage lies with the higher number of players.

##LANES

Each lane is filled with towers that deal abnormally high amounts of damage to the enemy, as well as lane minions that come in waves of 6 or 7 and are a simple aspect of the game that creates the most nuance out of any game mechanics. The towers only target the enemy if either of the following conditions are met: there are no enemy minions in the tower’s range, or an enemy is dealing damage to an allied player who is under the tower. In this way, even if you are a lot weaker than the enemy, the tower provides a lot of safety. Of course, this does not make a player invulnerable, as coordinated teams can cycle the tower’s targets by weaving in and out of tower range while trying to kill the player seeking protection from their tower. This is especially true at the highest levels of professional play where teams are extremely coordinated, which is the level this paper is investigating.

##LANE MINIONS

The lane minions work in interesting ways. Firstly, experience (exp) is gained if an enemy minion dies close to you. Secondly, if you are the player that deals the killing blow (last hit) to an enemy minion, you gain a small injection of gold. At first glance, the amount seems insignificant, but it adds up over time, and usually 20 more last hits than your lane opponent grants you a gold lead that is equivalent to killing them. Usually, the first concept beginner players looking to improve are told to work on is their last hitting. Not every minion is always worth last hitting, as sometimes, in the act of trying to last hit a minion, you expose yourself to deadly attacks from the enemy. A stronger enemy can use the threat of a deadly attack to prevent their lane opponent from getting last hits. In the top lane, it is also possible to deny your lane opponent exp by positioning very aggressively; however, this usually results in you not being able to secure all the last hits for yourself. The reason this doesn’t work in mid is due to the shape of the map, in which the effective length of the lane is much shorter as towers are close to each other. Since the top lane is longer, one mistake from an unsuspecting opponent can usually lead to a kill, whereas in the mid lane, they lose health but usually not enough to kill them before they return to the safety of their tower. The bot lane, while the same length as the top lane, isn’t usually able to zone the enemy from the exp range of lane minions as there are 2 players in each bot lane, so even with a strength lead, if you are less coordinated, you can lose the fight, and gold is usually worth more to a carry than exp. However, the bot lane has other options in which they can deny the enemy exp and gold. One way of doing this would be to stack a large minion wave and crash it under the tower. (No explanation of how to do this is provided as it is unnecessary, but essentially it means you have the lane minions from 3 or 4 different waves all under the enemy tower.) Then, they try to kill the enemy bot lane that is attempting to last hit the minions under the tower. Even if the tower kills you in the process, it is a net positive as you kill them so it evens out your death, but you also deny them 3-4 waves worth of gold and exp. Usually, when pro teams do this, they get assistance from their jungler to help kill under the tower. If they play it well and in the right situation, sometimes the enemy will die without trading back a kill.

##RESOURCES

There are three main ways that you can become stronger throughout the game. The first is through gold, which allows you to buy items once in your base. The second would be experience, which lets you level up after a certain threshold is reached. Each level gives you more stats (health, attack speed, armor, etc.) Finally, you can kill neutral objectives (an enemy to both teams), and whichever team secures the final blow to this objective unlocks a buff for their team. Gold and experience are straightforward to obtain. Last hitting minions was already discussed earlier and is the primary way of obtaining both gold and exp. Killing an enemy also grants a large amount of gold and experience, but it also puts them on a respawn timer in which they cannot last hit any minions, so they go further into a gold and exp deficit depending on the situation in which they died. Finally, killing enemy towers and neutral objectives also grant large amounts of gold. The neutral objectives spawn in increments and get progressively more powerful, but there are some stacking bonuses, so playing for these in the early game is a very viable strategy. Of these three, the objectives usually have the least impact in the early game and do not provide a great indicator as to whether a team is ahead, as many times teams will choose to trade an early neutral objective to secure more last hits, deny exp, or hit towers. As such, we do not consider the neutral objective statistics in our prediction of the match outcome.

##STRATEGY

These gold and experience advantages make your player (your champion) stronger, allowing you to win fights more easily and contest neutral objectives in the late game. However, if you play imperfectly, you can easily lose a fight even with a gold lead; for instance, if you miss your abilities. There is also the concept of “scaling,” in which some champions are a lot stronger in the late game but balanced by a weak early game. This sometimes means an early-game champion with a 60 cs lead (approximately 900 gold) may still be weaker than a late-game champion at 20 minutes into the game. A lot of newer players tend to overvalue the power of scaling, as even the most powerful late-game champions need to be executed with great precision to find success in the late game.

Another way a team at a gold deficit can win the game is by doing a “split push.” A split push, as the name suggests, is one that splits the map by pushing waves on both sides (the top lane and the bot lane). For instance, consider a scenario where one team (team A) is stronger than the other (team B) in a fight with all 10 members. However, team B has a top laner that is much stronger than any one of team A’s members. Team B can send their top laner to hit the enemy turrets in the top lane and their other 4 members to hit the enemy turrets in the bot lane. If team A only sends less than two players top lane, team B’s top laner can easily take team A’s top tower. If team A sends more than 1 player top lane, then team B can push the enemy bot lane towers as 4, and team A can only respond with 3. This favors team B in either scenario. While this is just a really surface-level analysis of a split push as there are many other factors to consider, it serves to show that through careful decision-making and precise play, the team that is behind on resources can still progress the game in their favor.
\newpage

##Implications of Findings

This paper analyzes the ability of T1 players and Gen G players to close out games after having certain leads. We consider two inputs: the cs at 10 minutes, which refers to how many enemy minions the player last hit before 10 minutes, and the highest level lead they had against their opponent. The maximum level lead helps the model predict a player's win rate of games in which they have a massive advantage over their opponent, as having a level lead of 2 or more is a significant advantage and usually means you also have a much higher cs lead. For this model, we do not consider the support players of either team, namely “Keria” and “Lehends,” as it is usually better for support players to let the carry last hit the enemy minions instead. Also, sometimes they can leave their carry to hit minions alone and surprise the enemy mid laner. This means they get less exp even though their team gets a net positive play. Also, levels and gold are best used on champions that deal damage as this is the best way to increase the damage output, whereas supports that give shields or lock down enemy champions do not benefit that much from gold or exp. In fact, support players usually try to let their teammates get the kills because gold is least effective in the hands of a support.

## Weaknesses and Next Steps

###Weakenesses

A major weakness of this analysis is that the the games under consideration were the solo queue games, with random teammates of a similar level. This version of league of legends has many differences to the professional play where team coordination is much more important than any individual player's skill level. These games were used as they are readily avaailable and generally speaking, the best professional players do really well in solo queue. Their solo queue characteristics are also apparent in their professional games. An example of this would Chovy, Gen G's mid laner. He is renowned for his extremely high cs and this shows up in both solo queue and professional matches. In this way, solo queue data are good approximations for analyzing certain aspects of player performance however the professional game data would be much more relevant for predicting professional games. Another weakness is the relatively small dataset. These players play thousands of games every year and we can analyze all of them to get a better understanding of how they play without cyclical shifts such as meta changes.

###Next Steps

A possible next step would be to obtain a product key from riot games which allows for more api requests and therefore we can analyze data from thousands of games at a time. Another potential next step would be to look at different input parameters as predictors of the match outcome. Maybe early neutral objectives as well as the type of champion selected as these both speak to the strength of a team at a certain point in time. These are all available on from the apis however there is no available data from a primary source about the professional games.
